---
title: 'Review of \texttt{R}'
author: "STA 360: Lab 1, Fall 2020 (This will not be graded or turned in)"
output: pdf_document
---

Today's agenda: A review of \texttt{R}, getting used to R markdown, vectors, matrices, scatterplots, and functions. 

***Lab Tasks***


1. Store three vectors using \textsf{rnorm()} of length $n = 100$ as \textsf{Var1}, \textsf{Var2}, and \textsf{Var3}.

```{r}
set.seed(1)
Var1 <- rnorm(100)
set.seed(2)
Var2 <- rnorm(100)
set.seed(3)
Var3 <- rnorm(100)
```



2. List all the items currently in the environment.
```{r}
ls()
```


3. Store \textsf{Var1} in a $10 \times 10$ matrix. Call this \textsf{myMatrix}.
```{r}
myMatrix <- matrix(Var1, 10, 10)
```


4. Create a scatterplot of \textsf{Var1} vs. \textsf{Var2}. On the same plotting window include histograms of \textsf{Var1} and \textsf{Var2}.
```{r}
par(mfrow=c(1,3))
plot(Var1, Var2)
hist(Var1)
hist(Var2)
```


5. Write a function that takes as its inputs, $p = 2$, $n$-dimensional vectors and a vector of length $p$ containing the names of these vectors. Your function combine these two vectors into a \textsf{data.frame()}, get the row-wise maximum and store this in a new vector. Finally produce a box-plot of this vector, store it as a separate .pdf, and return the mean value of this vector.

```{r}
func <- function(v1, v2, myNames) {
  myData = data.frame(v1, v2)
  names(myData) = myNames
  myVector = apply(myData, 1, max)
  pdf('myPlot.pdf')
  boxplot(myVector)
  dev.off()
  final <- mean(myVector)
  return(final)
}

func(Var1, Var2, c('myVar1', 'myVar2'))
```


